### 2017-07-09 background.tex
\it{DATA}
comments

A tuple $(t_i, p_i)$ in the packet trace means the packet $p_i$ is presented to
the state machine at time $t_i$. The monitor {\it rejects a trace} $Tr$ if there
exists a prefix of $Tr$ such that all states reachable after consuming the
prefix have no valid transitions for the next $(t, p)$ input.

As an example, the monitor state machine illustrated in
Fig.~\ref{fig:dot11_tx_ta} can be formally defined as follows:

\begin{itemize}
  \item $\Sigma = \{DATA_i, DATA'_i, Ack\ |\ 0 \le i < N\}$.
  \item Clock variables $C = \{c\}$. The only clock variable $c$ is
    used for acknowledgment time out.
  \item $\mathbb{X} = \{i\}$, as a variable with ${\mathit log}(N) + 1$ bits to
    count from $0$ to $N$.  
  \item Guard constraints $G = \{ c \le T_o, c > T_o, T_o < c \le T_m\}$.
    $T_o$ is the acknowledgment time out value, and $T_m >
    T_o$ is the maximum delay allowed before the retransmission packet gets
    sent. $T_o$ can be arbitrary large but not infinity in order to check the
    liveness of the DUT.
\end{itemize}

>>>>

A tuple $(t_i, p_i)$ in the packet trace means the packet $p_i$ is presented to
the state machine at time $t_i$.  As an example, the monitor state machine
illustrated in Fig.~\ref{fig:dot11_tx_ta} can be formally defined as follows:

\begin{itemize}
  \item $\Sigma = \{\it{DATA}_i, \it{DATA}'_i, Ack\ |\ 0 \le i < N\}$.
  \item Clock variables $C = \{c\}$. The only clock variable $c$ is
    used for acknowledgment time out.
  \item $\mathbb{X} = \{i\}$, as a variable with ${\mathit log}(N) + 1$ bits to
    count from $0$ to $N$.  
  \item Guard constraints $G = \{ c \le T_o, c > T_o, T_o < c \le T_m\}$.
    $T_o$ is the acknowledgment time out value, and $T_m >
    T_o$ is the maximum delay allowed before the retransmission packet gets
    sent. In this particular case, $T_o$ can be arbitrary large but not infinity
        in order to check the liveness of the DUT.
\end{itemize}

\begin{definition}[Trace Rejection and Acceptance]
    A monitor state machine $S$ {\it rejects a trace} $Tr$ if there
    exists a prefix of $Tr$ such that all states reachable after consuming the
    prefix have no valid transitions for the next $(t, p)$ input. Otherwise, $S$
    {\it accepts} $Tr$ (or $Tr$ {\it satisfies} $S$).
\end{definition}

### 2017-07-09 case/tx.tex
comments

@@

The different between two packet traces can be quantified by the Jaccard distance
metric.

@@

A na\"ive way to calculate the Jaccard distance of two traces is to use the hash
of the $(time, packet)$ pair for set intersection and union operation.  However,
it does not work for mutation trace which contains fabricated packets with no
actual payload.  Therefore, we use a protocol specific canonical representation
of packets when calculating the distance.  In particular, the string
\texttt{r\_\it{DATA}\_i\_t} represents the $t^{th}$ transmission of a data packet
with sequence number $i$, and $r$ represents the round of sequence numbers as it
wraps after 4096.  And similarly \texttt{r\_Ack\_i\_t} is the corresponding
acknowledgment packet.

@@

We then evaluate the cost of resolving uncertainty. In particular, we use
the Average Search Steps Per Packet (ASSPP) as a metric to quantify the search
cost.  It is calculated by dividing the total number of search steps by the
number of packets in the packet trace. For DUT traces, ASSPP is always 1 since
there is no uncertainty. For sniffer traces, however, multiple search steps must
be to conducted to resolve the potential uncertainty of each packet in the
sniffer trace.

>>>>

The DUT and endpoint are configured to use the IEEE 802.11g standard with both
RTS/CTS and fragmentation disabled.  A Constant Bit Rate (CBR) UDP traffic
(54~Mbps) is generated from the DUT to the endpoint.  The UDP packet size is
1436 bytes, which results in a 1500 bytes \wifi{} packet.

In order to control the packet loss ratios between each pair of devices, we
developed a new propagation loss model for \ns{} called
\texttt{MatrixRandomPropagationLossModel}.
\sloppy{%
  Instead of a constant propagation
  loss as in existing \texttt{MatrixPropagationLossModel}, the signal
  propagation loss between a pair of nodes is determined by a binary random
  variable of two  values: 0~dB (no loss) and 1000~dB (complete loss).
  }
Therefore, any packet loss
probability can be achieved by adjusting the random variable distribution.
The
model supports both symmetric and asymmetric propagation losses.
We use
symmetric propagation loss in all our experiments.
Finally, pcap capture is enabled in both the DUT and the sniffer devices.

@@

The difference between two packet traces can be quantified by the Jaccard distance
metric.

@@

A na\"ive way to calculate the Jaccard distance of two traces is to use the hash
of the $(time, packet)$ pair for set intersection and union operation.  However,
it does not work for mutation trace which contains fabricated packets with no
actual payload.  Therefore, we use a protocol specific canonical representation
of packets when calculating the distance.  In particular, the string
\textit{r\_DATA\_i\_t} represents the $t^{th}$ transmission of a data packet
with sequence number $i$, and $r$ represents the round of sequence numbers as it
wraps after 4096.  And similarly \textit{r\_ACK\_i\_t} is the corresponding
acknowledgment packet.

@@

We then evaluate the cost of resolving uncertainty. In particular, we use
the Average Search Steps Per Packet (ASSPP) as a metric to quantify the search
cost.  It is calculated by dividing the total number of search steps by the
number of packets in the packet trace. For DUT traces, ASSPP is always 1 since
there is no uncertainty. For sniffer traces, however, multiple search steps must
be conducted to resolve the potential uncertainty of each packet in the
sniffer trace.

### 2017-07-09 case/xbox.tex
comments

\begin{table}[t!]
  \centering
  \begin{tabular}{lll}
    \toprule
    \textbf{Protocol Aspects} & \textbf{Traces} & \textbf{Violations (\%)}\\
    \midrule
    Sequence Number & 3049 & 1539 (50.48\%) \\
    Station Scheduling & 3046 & 2045 (67.14\%) \\
    Uplink Modulation & 3127 & 8 (0.26\%) \\
    Downlink Modulation & 3127 & 24 (0.77\%) \\
    \bottomrule
  \end{tabular}
  \caption{\textbf{Validation Results on Traces from the Gaming Wireless Protocol.}}
  \label{tab:xbox}
\end{table}

@@

We split the traces into \num{100000} packet segments, and applied our framework
to validate the DUT implementation. We found that the latest implementation of
the protocol under development had violations of the protocol specification.
Some of the implementation bugs we found related to the sequence number
management, station scheduling during power saving mode, and modulation rates
adaptation.

Table~\ref{tab:xbox} summarizes the validation result. Note that if we disable
the augmented transitions, each trace will be flagged as violation because of
the missing packets, thereby reducing the usability of the tool. We also note
that some bugs manifest more often than others. For instance, the bugs related
to packet sequence number and station scheduler were detected in about half the
traces, and the bug related to the rate control algorithm was detected in only a
few traces. This is because the previous two aspects are essential in all
protocol operations, while the bugs related to rate control only manifest
themselves under certain link conditions.

After communicating with the testing team, we confirmed that the sequence number
bug was already known, as it is relatively easy to detect even by manually
examining the traces. The bug related to station scheduling was also noticed
before, yet no quantitative results about how frequent this bug appears were
obtained because of the lack of automatized validating tools. Finally, the bug
related to rate control has been filed as a bug report. All reported bugs were
fixed before the next release the product.

>>>>

\begin{table}[t!]
  \centering
  \begin{tabular}{lll}
    \toprule
    \textbf{Protocol Aspects} & \textbf{Traces} & \textbf{Violations (\%)}\\
    \midrule
    Sequence Number & 3049 & 1539 (50.48\%) \\
    Station Scheduling & 3046 & 2045 (67.14\%) \\
    Uplink Modulation & 3127 & 8 (0.26\%) \\
    Downlink Modulation & 3127 & 24 (0.77\%) \\
    \bottomrule
  \end{tabular}
  \caption{\textbf{Validation Results on Traces from the Gaming Controller Wireless Protocol.}}
  \label{tab:xbox}
\end{table}

@@

We first split the traces into \num{100000} packet segments, which yields \num{3127}
traces for testing.  We then applied our framework on the traces to validate the
DUT implementation. We found that the latest implementation of the protocol
under development had violations of the protocol specification.  Some of the
implementation bugs we found related to the sequence number management, station
scheduling during power saving mode, and modulation rates adaptation.
Table~\ref{tab:xbox} summarizes the validation results. The \textbf{Traces}
column shows the number of traces that we have successfully validated for each
category, and the third column shows the fraction of traces that contain at
least one violation in that category.

Note that if we disable
the augmented transitions, each trace will be flagged as violation because of
the missing packets, thereby reducing the usability of the tool. We also note
that some bugs manifest more often than others. For instance, the bugs related
to packet sequence number and station scheduler were detected in about half the
traces, and the bug related to the rate control algorithm was detected in only a
few traces. This is because the previous two aspects are essential in all
protocol operations, while the bugs related to rate control only manifest
themselves under certain link conditions.

After communicating with the testing team, we confirmed that the sequence number
bug was already known, as it is relatively easy to detect even by manually
examining the traces. The bug related to station scheduling was also noticed
before, yet no quantitative results about how frequent this bug appears were
obtained because of the lack of automatized validating tools. Finally, the bug
related to rate control, which was unknown previously, has been filed as a bug
report. All reported bugs were fixed before the next release the product.

### 2017-07-09 framework/augment.tex
all-in-one
comments

\begin{figure}[t!]
  \centering
  \includegraphics[width=0.7\textwidth]{./figures/dot11_tx_checker.pdf}
  \caption{\textbf{Augmented Monitor State Machine.} Augmented transitions are
  highlighted in bold face. $\overline{Pkt}$ means either $\epsilon$ or $Pkt$.}
  \label{fig:augment}
\end{figure}

@@

A accepting run of $S^+$ on sniffer trace $Tr$ yields a mutation trace $Tr'$
which represents one possibility of the DUT trace. Specifically, $Tr'$ can be
obtained by adding missing packets indicated by Type-1 transitions to $Tr$, and
removing extra packets indicated by Type-2 transitions from $Tr$

@@

\begin{theorem}
  There exists a mutation trace $Tr' \in \mathcal{M}(Tr)$ that satisfies $S$ if
  and only if $Tr$ satisfies $S^+$.
 \label{the:equivalent}
\end{theorem}
\begin{proof}
  Assume $Tr$ satisfies $S^+$, and $P$ is a sequence of accepting transitions,
  we construct a mutation trace $Tr'$ using $P$ and show that $Tr'$ satisfies
  $S$.

@@

  Now we show that there exists a accepting transition sequence $P'$ of $S^+$ on
  input $Tr'$ that does not contain augmented transitions.  In particular, $P'$
  can be obtained by substituting all Type-1 transitions with corresponding
  original transitions, and removing all Type-2 transition.  Since $P'$ does not
  contain augmented transitions, it is also an accepting transition sequence of
  $S$ on input $Tr'$, hence $Tr'$ satisfies $S$.

@@

  We construct a accepting transition sequence $P$ of $S^+$ on input $Tr$ as
  follows.
  \begin{itemize}
    \item For each packet $p \in Tr' \setminus Tr$, substitute the transition
      $\langle s_i, v_i, s_j, v_j, p, g, C'\rangle$ with the corresponding Type-1
      transition $\langle s_i, v_i, s_j, v_j, \epsilon, g, C'\rangle$.
    \item For each transition $\langle s_i, v_i, s_j, v_j, \sigma, g, C'\rangle$
      followed by packet $p \in Tr\setminus Tr'$, add a Type-2 self
      transition $\langle s_j, v_j, s_j, v_j, p, g, \emptyset\rangle$. This is
      possible since $Tr'$ is a mutation trace of $Tr$, thus  for all $p \in Tr'
      \setminus Tr$, $p.src \ne DUT$.
  \end{itemize}
  Therefore, $Tr$ satisfies $S^+$.
  \qed
\end{proof}

>>>>

\begin{figure}[t!]
  \centering
  \includegraphics[width=0.7\textwidth]{dot11_tx_checker.pdf}
  \caption{\textbf{Augmented Monitor State Machine.} Augmented transitions are
  highlighted in bold face. $\overline{Pkt}$ means either $\epsilon$ or $Pkt$.}
  \label{fig:augment}
\end{figure}

@@

An accepting run of $S^+$ on sniffer trace $Tr$ yields a mutation trace $Tr'$
which represents one possibility of the DUT trace. Specifically, $Tr'$ can be
obtained by adding missing packets indicated by Type-1 transitions to $Tr$, and
removing extra packets indicated by Type-2 transitions from $Tr$

@@

\begin{theorem}
    Let $Tr$ be a sniffer trace, $\mathcal{M}(Tr)$ be $Tr$'s mutation traces, $S$
    and $S^+$ be the original and augmented monitor state machine respectively.
  There exists a mutation trace $Tr' \in \mathcal{M}(Tr)$ that satisfies $S$ if
  and only if $Tr$ satisfies $S^+$.
 \label{the:equivalent}
\end{theorem}
\begin{proof}
  Assume $Tr$ satisfies $S^+$, and $P$ is a sequence of accepting transitions,
  we construct a mutation trace $Tr'$ using $P$ and show that $Tr'$ satisfies
  $S$.

@@

  Now we show that there exists an accepting transition sequence $P'$ of $S^+$ on
  input $Tr'$ that does not contain augmented transitions.  In particular, $P'$
  can be obtained by substituting all Type-1 transitions with corresponding
  original transitions, and removing all Type-2 transitions.  Since $P'$ does not
  contain augmented transitions, it is also an accepting transition sequence of
  $S$ on input $Tr'$, hence $Tr'$ satisfies $S$.

@@

  We construct an accepting transition sequence $P$ of $S^+$ on input $Tr$ as
  follows.
  \begin{itemize}
    \item For each packet $p \in Tr' \setminus Tr$, substitute the transition
      $\langle s_i, v_i, s_j, v_j, p, g, C'\rangle$ with the corresponding Type-1
      transition $\langle s_i, v_i, s_j, v_j, \epsilon, g, C'\rangle$.
    \item For each transition $\langle s_i, v_i, s_j, v_j, \sigma, g, C'\rangle$
      followed by packet $p \in Tr\setminus Tr'$, add a Type-2 self
      transition $\langle s_j, v_j, s_j, v_j, p, g, \emptyset\rangle$. This is
      possible since $Tr'$ is a mutation trace of $Tr$, thus  for all $p \in Tr'
      \setminus Tr$, $p.src \ne DUT$.
  \end{itemize}
  Therefore, $Tr$ satisfies $S^+$.
  \qed
\end{proof}

### 2017-07-09 framework/search.tex
comments

The main routines of the algorithm are shown in Algorithm~\ref{alg:search}.  In
the top level \texttt{SEARCH} routine, we first obtain the augmented state
machine $S^+$, and then call the recursive \texttt{EXTEND} function with an empty
prefix, the sniffer trace, and the $S^+$'s initial state. 
In the \texttt{EXTEND} function, we try to consume the first packet in the
remaining trace using either Type-0, Type-1 or Type-2 transition.
Note that we always try to use Type-0 transitions before other two augmented
transitions (line~\ref{alg:search:order}).
This ensures the first found mutation trace will have the most number of Type-0
transitions among all possible mutation traces.
Intuitively, this means the search algorithm tries to utilize the sniffer's
observation as much as possible before being forced to make assumptions.

>>>>

The main routines of the algorithm are shown in Algorithm~\ref{alg:search}.  In
the top level \texttt{SEARCH} routine, we first obtain the augmented state
machine $S^+$, and then call the recursive \texttt{EXTEND} function with an empty
prefix, the sniffer trace, and the $S^+$'s initial state. 
In the \texttt{EXTEND} function, we try to consume the first packet in the
remaining trace using either Type-0, Type-1 or Type-2 transition.
Note that we always try to use Type-0 transitions before other two augmented
transitions (line~\ref{alg:search:order}).
This ensures the first found mutation trace will have the most number of Type-0
transitions among all possible mutation traces.
Intuitively, this means the search algorithm tries to utilize the sniffer's
observation as much as possible before considering the packet loss by the
sniffer or DUT.
